---
title: "Tutorial: Time Profile Plotting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial_Timeprofiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  eval = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.width = 7,
  comment = "#>"
)
```

```{r setup, echo=FALSE, eval=TRUE}
library(ospsuite.reportingframework)
library(data.table)
options(rmarkdown.html_vignette.check_title = FALSE)
```


```{r loadProject, echo=FALSE, eval=TRUE}
ggplot2::theme_set(theme_osp())
options(knitr.kable.NA = '')

runPlotTheme <- ggplot2::theme(
  legend.position = "top",
  legend.title = ggplot2::element_blank()
)


rootDirectory <- file.path(tempdir(), "testProject")
configurationDirectory <- file.path(rootDirectory, "Scripts", "ReportingFramework")

initProject(
  projectDirectory = rootDirectory,
  configurationDirectory = configurationDirectory
)

projectConfiguration <- ospsuite.reportingframework::createProjectConfiguration(
  path = file.path(configurationDirectory, "ProjectConfiguration.xlsx")
)

instDirectory <- system.file(
  package = "ospsuite.reportingframework",
  "extdata",
  mustWork = TRUE
)

configFiles <- list.files(file.path(instDirectory, "Tutorial"), pattern = ".xlsx", full.names = TRUE)
invisible(lapply(configFiles, function(f) {
  file.copy(
    from = f,
    to = file.path(projectConfiguration$configurationsFolder, basename(f)), overwrite = TRUE
  )
}))

dir.create(file.path(rootDirectory, "Data"))
dataFiles <- list.files(file.path(instDirectory, "Tutorial"), pattern = ".csv", full.names = TRUE)
invisible(lapply(dataFiles, function(f) {
  file.copy(
    from = f,
    to = file.path(rootDirectory, "Data", basename(f)), overwrite = TRUE
  )
}))

invisible(file.copy(
  from = file.path(instDirectory, "Populations", "adults.csv"),
  to = file.path(projectConfiguration$populationsFolder, "random_population.csv"), overwrite = TRUE
))

invisible(file.copy(
  from = file.path(instDirectory, "Tutorial", "study_1234_population.csv"),
  to = file.path(projectConfiguration$populationsFolder, "study_1234_population.csv"), overwrite = TRUE
))

modelFiles <- list.files(file.path(instDirectory, "Models"), full.names = TRUE)
invisible(lapply(modelFiles, function(f) {
  file.copy(from = f, to = file.path(projectConfiguration$modelFolder, basename(f)), overwrite = TRUE)
}))

scenarioList <-
  createScenariosWrapped(
    projectConfiguration = projectConfiguration,
    scenarioNames = NULL
  )

scenarioResults <- runAndSaveScenarios(
  projectConfiguration = projectConfiguration,
  scenarioList = scenarioList
)

dataObserved <- readObservedDataByDictionary(projectConfiguration = projectConfiguration)

# add aggregated  groups of data
dataObserved <- rbind(dataObserved,
  aggregateObservedDataGroups(
    dataObserved = dataObserved,
    groups = c("1234_iv", "1234_po")
  ),
  fill = TRUE
)
```

## Introduction

This guide walks you through generating time profile plots using the `{ospsuite.reportingframework}` package, step by step from a single simulated individual through to a fully assembled Word report.

### Example Study

Throughout the tutorial, we work with a synthetic crossover pharmacokinetic study (Study 1234) in which a small molecule was administered to multiple subjects via two routes:

- **IV**: a single 1 mg intravenous infusion over 5 minutes (`iv_1_mg_5_min.pkml`)
- **PO**: a single 3 mg oral solution (`po_3_mg_solution.pkml`)

Because this is a crossover design, the same subjects appear in both the IV and PO groups. Observed concentration–time data are provided in `timeprofiles_study1234_iv.csv` and `timeprofiles_study1234_po.csv`.

Each section of the tutorial adds one layer of complexity — virtual twin populations, random populations, a reference scenario, custom time-range windows — and ends with packaging everything into a Word report. In each section, the rendered configuration table shows exactly what to edit in Excel before running the corresponding `runPlot()` call.

## Prerequisites
Before you begin, ensure you have installed the `ospsuite.reportingframework` package. Familiarity with R, RStudio, and basic Excel operations will help you follow the steps.

## Package Materials

The package contains the following materials used for the tutorial:

- Data files used for the tutorial: `timeprofiles_study1234_iv.csv`, `timeprofiles_study1234_po.csv`
- Final configuration tables: `DataImportConfiguration.xlsx`, `Individuals.xlsx`, `Populations.xlsx`, `Scenarios.xlsx`, `Reports.xlsx`
- The final workflow: `workflow.R`
- The final Word document containing all plots generated during this tutorial: `TutorialPlots.docx`

You can access the folder containing this material using the command below:

```{r tutorial-path, eval = FALSE}
print(system.file(
  "extdata", "Tutorial",
  package = "ospsuite.reportingframework",
  mustWork = TRUE
))
```
You can access the folder containing the models used in the tutorial ("iv_1_mg_5_min.pkml" and "po_3_mg_solution.pkml") using the command below:

```{r model-path, eval = TRUE}
print(system.file(
  "extdata", "Models",
  package = "ospsuite.reportingframework",
  mustWork = TRUE
))
```

## Table of Contents

1. [Setting Up the Project and Basic Simulations](#1-setting-up-the-project-and-basic-simulations)  
   Learn how to set up your project, save a template workflow, and initialize the project structure for time profile plotting.

2. [Adding virtual twin populations](#2-adding-virtual-twin-populations)  
   Discover how to define and export virtual twin populations for more refined simulations.

3. [Adding Random Populations and Use of Aggregated Data](#3-adding-random-populations-and-use-of-aggregated-data)  
   Understand how to aggregate individual data and work with aggregated random populations for comparative analysis.

4. [Adding Reference Scenarios](#4-adding-reference-scenarios)  
   Learn how to overlay a reference scenario on a plot to compare two conditions side by side.

5. [Adding plots with different time ranges](#5-adding-plots-with-different-time-ranges)  
   Explore how to visualize pharmacokinetic data across various time ranges to reveal different drug behavior aspects.

6. [Report Creation](#6-report-creation)  
   Learn how to merge plots into a report and render it as a Word document.

## 1) Setting Up the Project and Basic Simulations

### Step 1: Save Template Workflow

- Select your project directory (`<rootDirectory>`). In this directory, the reporting framework
folder structure will be established (if not already existing).
- Verify or create the subfolder `<rootDirectory>/Scripts/ReportingFramework` (referred to as `<configurationDirectory>` throughout this tutorial).
- Open RStudio and navigate to the Addins menu.
- Select "Open Template Workflow" to create a new workflow file. Save the template in the `<configurationDirectory>` as `workflow.R`. At the top of the workflow file, you can add a description of the purpose of your workflow.

### Step 2: Initialize Project Structure

Execute the first lines (listed below) of your workflow. This will:

- Change the working directory to the workflow file location.
- Set some graphic defaults.
- Set some options to switch between final valid runs and preliminary runs, such as watermarks.
- Initialize the project directory.
- Create or update `.gitignore` files in configuration folders so `*.xlsx` files are ignored by Git.
- Load the project configuration.
- Set up a logfile to catch messages.

```{r print-template_workflow_initialization}
# Initialization  ----------------------------------------------------------
# Load necessary libraries and source project-specific code
library(ospsuite.reportingframework)

# Set graphic defaults
# (see vignette(package = 'ospsuite.plots', topic = 'ospsuite_plots'))
ggplot2::theme_set(theme_osp())
options(knitr.kable.NA = "")

runPlotTheme <- ggplot2::theme(
  legend.position = "top",
  legend.title = ggplot2::element_blank()
)

# Set this to TRUE if you want to execute the workflow as a final valid run.
# (see ?setWorkflowOptions)
setWorkflowOptions(isValidRun = FALSE)

# Setup project structure -------------------------------------------------
# Create project directory and initialize the structure
# (see ?initProject)
initProject()

# Get paths of all relevant project files and folders
projectConfiguration <- ospsuite.reportingframework::createProjectConfiguration(
  path = file.path("Scripts", "ReportingFramework", "ProjectConfiguration.xlsx")
)
# For older ProjectConfiguration.xlsx files without a stored framework version,
# use ignoreVersionCheck = TRUE (see ?createProjectConfiguration).

# Initialize log file
initLogfunction(projectConfiguration)
```


When loading an existing project configuration that was created before version tracking was introduced, pass `ignoreVersionCheck = TRUE` to `createProjectConfiguration()` or update the file once with `projectConfiguration$save()`.

### Configuration in Git: Save JSON Snapshot

Configuration Excel files are intended as editable working files and are ignored by Git in configuration folders. The shareable/versioned representation is `ProjectConfiguration.json`.

After changing configuration tables, create or refresh the JSON snapshot:

```{r save-config-json, eval = FALSE}
projectConfiguration <- ospsuite.reportingframework::createProjectConfiguration(
  path = file.path(configurationDirectory, "ProjectConfiguration.xlsx")
)

ospsuite.reportingframework::snapshotProjectConfigurationRF(
  projectConfiguration,
  outputDir = configurationDirectory
)
```

Commit `ProjectConfiguration.json` to Git.

To recreate the configuration from JSON:

```{r restore-config-json, eval = FALSE}
projectConfiguration <- ospsuite.reportingframework::restoreProjectConfigurationRF(
  snapshotFilePath = file.path(configurationDirectory, "ProjectConfiguration.json"),
  outputDir = configurationDirectory
)
```

### Step 3: Copy the Sample Data to Your Project Structure

You have now the complete project structure with all configuration files. Copy the model files to the folder `<rootDirectory>/Models` and the data files to `<rootDirectory>/Data`.


### Step 4: Read Observed Data

(See **[Data Import by Dictionary](DataImport_by_Dictionary.html)**)

Open the `<configurationDirectory>/DataImportConfiguration.xlsx`.


- List in the sheet `DataFiles` your available data. The paths should be relative to the configuration table which is in `<rootDirectory>/Scripts/ReportingFramework`, the data are saved in `<rootDirectory>/Data`, so the correct path is `../../Data`.
- Each file gets a file identifier, which can be used as a filter in the `readObservedDataByDictionary` function.
- As dictionary, select `tpDictionary`.
- We do not need a filter, so ensure the column `DataFilter` is empty.
- The data is of class `tp Individual` (individual time profile data).


```{r configData-Files, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$dataImporterConfigurationFile,
  sheetName = "DataFiles",
  skipDescriptionRow = TRUE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp[DataClass == DATACLASS$tpIndividual])
```

- Match existing columns of the datasets when available.
- For `individualId` and `group`, concatenate existing columns using the columns `Filter` and `FilterValue`.
- For the 'time profile' columns, use the source columns `Time`, `DV`, `DVUNIT`, and `LLOQ`.
- For `biometrics` columns, use the source columns `Age`, `WGHT0`, `HGHT0`, and `SEX`.
- Set `population` to `European_ICRP_2002` by the filter option.
- Metadata `route` is available; `dose` and `outputPathId` are set via the filter option, each occupying a separate row in the dictionary.


```{r configData-Dictionary, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$dataImporterConfigurationFile,
  sheetName = "tpDictionary",
  skipDescriptionRow = TRUE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp)
```

Now you can execute the next workflow line:

```{r readData}
# Read data as data.table
dataObserved <- readObservedDataByDictionary(projectConfiguration = projectConfiguration)
```

This will return a data.table with the observed data and add data to the configuration tables.

- The biometrics of the individuals in the observed data are added to the `Individuals.xlsx`. Information not in the observed data, like ontogeny, must be added manually.


```{r individualbiometrics, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$individualsFile,
  sheetName = "IndividualBiometrics",
  skipDescriptionRow = FALSE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp)
```

- The identifiers for the data groups and outputs are added to the `Reports.xlsx`, sheets `DataGroups` and `Outputs`. These tables configure properties for output plots such as `DisplayName` and `DisplayUnit`. Please add the `OutputPaths` for the `OutputPathsIds`.


```{r datagroups, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$addOns$reportsFile,
  sheetName = "Outputs",
  skipDescriptionRow = TRUE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp)
```


### Step 5: Set Up and Run Scenario

We will skip the section on exporting populations and start with a simple scenario to simulate one individual, defined by biometrics, with an iv application.

Open the `Scenarios.xlsx` and edit the sheet `Scenarios`. Fill in the columns:

- `Scenario_name`: identifier of the scenario.
- `IndividualId`: corresponds to the individual identifier in the configuration table `Individuals.xlsx`.
- `ModelFile`: this file must exist in the folder `Models`


```{r scenario_individual, echo=FALSE, eval = TRUE}
dtScenarios <- xlsxReadData(
  wb = projectConfiguration$scenariosFile,
  sheetName = "Scenarios",
  skipDescriptionRow = FALSE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

setcolorder(dtScenarios, neworder = "ModelFile", after = "ReadPopulationFromCSV")

knitr::kable(dtScenarios[1])
```

Now we can set up the `scenarioList` and run the simulation:


```{r setupsimulation}
scenarioList <-
  createScenariosWrapped(
    projectConfiguration = projectConfiguration,
    scenarioNames = NULL
  )

scenarioResults <- runAndSaveScenarios(
  projectConfiguration = projectConfiguration,
  scenarioList = scenarioList,
  simulationRunOptions = ospsuite::SimulationRunOptions$new(
    showProgress = TRUE
  )
)
```

The results are saved in `<rootDirectory>/Outputs/ReportingFramework/SimulationResults`.

### Step 6: Configure and Generate Plots

For the plot configuration, call a helper function that adds all defined scenarios with default settings to the plot configuration table. This helper function changes the input configuration of a workflow and is not intended to be part of a final workflow.

```{r addDefaultConfigForTimeProfilePlots}
addDefaultConfigForTimeProfilePlots(
  projectConfiguration = projectConfiguration,
  sheetName = "TimeProfiles",
  overwrite = FALSE
)
```

Then we open the `Reports.xlsx` sheet `TimeProfiles`. It starts with a header line and one line for the previously defined scenario.

- In the `Section` column, add a level-2 header row (e.g., `## 1) Basic Individual Simulation`). This header will appear as a section title in the exported Word report.

- Edit the columns:
  - `PlotName`: The plot name should be meaningful as it is also used for the exported figure names.
  - `PlotCaptionAddon`: Add-on for the caption used for all scenarios in this plot.
  - `DataGroupIds`: Data group identifier corresponding to this scenario, must be available in `dataObserved$group`.
  - `IndividualIds`: Identifier of the individual, should correspond to the `IndividualId` of the scenario and the `individualId` in the observed data.
  - `TimeRange_firstApplication` and `TimeRange_lastApplication`: Set to empty, as we have a single dose.

This line will create a time-profile plot that compares the simulated data to the observed data.

We also want to add goodness of fit plots. Because the two outputs have different physical dimensions (e.g., plasma concentration and AUC), goodness of fit plots must be restricted to a single dimension at a time to keep the axes consistent. We do this in a separate row:

- Copy the plot line and edit the columns:
  - `PlotName`: each plot must have a unique Plot identifier.
  - `OutputPathIds`: ensure we have only outputs of one dimension selected.
  - Deselect `Plot_TimeProfiles`.
  - Select `Plot_PredictedVsObserved`, `Plot_ResidualsAsHistogram`, `Plot_ResidualsVsTime`, `Plot_ResidualsVsObserved`, and `Plot_QQ`.

Check also the sheet `Scenarios`. Here you can define for each scenario a `LongName` that will be used in captions as scenario descriptions.

```{r plotconfig-1, echo=FALSE, eval = TRUE}
dtPlots <- xlsxReadData(
  projectConfiguration$addOns$reportsFile,
  sheetName = "TimeProfiles",
  skipDescriptionRow = TRUE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(dtPlots[1:4]
 |>  dplyr::select(!any_of(c("TimeRange_h0_6", "TimeRange_h6_24"))))
```

Now we can execute the `runPlot` call to generate the plots. For now, we suppress the export of the figures by setting the input parameter `plotNames = c('one_individual', 'one_individual-gof')`. `runPlot` returns a list of ggplot objects. The caption is added as an attribute to each plot.


```{r plot1, eval = TRUE}
plotList <- runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  plotNames = c("one_individual", "one_individual-gof"),
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults
  )
)
```

Below the logarithmic time-profile is displayed. See the Word document for the other plots.

```{r plot1figure, eval = TRUE,fig.height = 2.8, results='asis', fig.alt="Logarithmic time-profile plot comparing simulated data to observed data."}
print(plotList[["one_individual-TP-log-total"]])
print(attr(plotList[["one_individual-TP-log-total"]], "caption"))
```

## 2) Adding Virtual Twin Populations

(See **[Population](Population.html)**).

### Step 1: Edit Population Configuration

Manually adjust your Excel configuration to define virtual twin populations. The `VirtualTwinPopulation` sheet within the `Individuals.xlsx` file was created during data import. It contains all data groups with individual data. Since this is a crossover study and the individuals in group 'IV' are also in group 'PO', only one population is suggested for both data groups.

For now, we will leave the default configuration and only adjust the `populationName`:


```{r configure-virtual-population, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$individualsFile,
  sheetName = "VirtualTwinPopulation",
  skipDescriptionRow = FALSE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp)
```

### Step 2: Export Population

- Go to the Export populations section of the workflow and uncomment the lines for the virtualTwin export.
- Adjust the input parameter `modelFile`. The model file is used for unit conversions. Because the virtual twin population contains only individual physiological parameters (height, weight, age, etc.) and no dosing-protocol parameters, either model file can be used.

Now you can execute the following lines:

```{r export-virtual-population, eval=FALSE}
exportVirtualTwinPopulations(
  projectConfiguration = projectConfiguration,
  populationNames = NULL,
  modelFile = "po_3_mg_solution.pkml",
  overwrite = FALSE
)
```

This will create a population based on the configuration table from Step 1. It is saved as a CSV file in `Models/Populations`.

### Step 3: Set Up and Run Scenario

Open `Scenarios.xlsx` again and add a new row to the sheet `Scenarios`. Instead of `IndividualId`, fill `PopulationId` with the name of the newly created population and set the value to `1` (TRUE) in the column `ReadPopulationFromCSV`.

```{r scenario-virtual-twin, echo=FALSE, eval = TRUE}
knitr::kable(dtScenarios[1:2], align = "c")
```

To omit a rerun from previously defined scenarios, we change the `scenarioNames` parameter in the `scenarioList` initialization from `NULL` to the names of the newly created scenarios.


```{r simulate-virtual-twin, eval=FALSE}
scenarioList <-
  createScenariosWrapped(
    projectConfiguration = projectConfiguration,
    scenarioNames = dtScenarios$Scenario_name[2]
  )

scenarioResults <- runAndSaveScenarios(
  projectConfiguration = projectConfiguration,
  scenarioList = scenarioList,
  simulationRunOptions = ospsuite::SimulationRunOptions$new(
    numberOfCores = NULL,
    checkForNegativeValues = NULL,
    showProgress = TRUE
  )
)
```

### Step 4: Configure and Generate Plots

We will add the new scenario to the plot configuration table using the helper function `addDefaultConfigForTimeProfilePlots`.

Then we open the configuration table and adjust the plot configurations.

First, we add a plot as an example to filter individuals explicitly:

- Copy the `virtual_twin_population` line.
- Adjust the column `plotName` to generate two independent plots.
- Add the `DataGroupId` and as `IndividualId` for the first plot the first three IDs of the data set, and for the second plot the last three.
- Set a bracket around the outputs in the column `OutputPathIds`. This way, both outputs will be plotted in one panel.

Then we add a plot as an example to filter individuals with shortcuts:

- Copy one of the previous plots for the `virtual_twin_population` scenario.
- As `IndividualId`, use `'*'`. Now all available individuals will be plotted. If you have many individuals, this could lead to very large plots with many rows, so the plot is split into different plots. The maximum row number is an input parameter of the plot function.

As we did in Section 1, we will add a line for the goodness of fit plots. Copy the plot line for the last plot and edit the columns:

- `PlotName`
- `OutputPathIds`: ensure we have only outputs of one dimension selected.
- Deselect `Plot_TimeProfiles`.
- Select `Plot_PredictedVsObserved`, `Plot_ResidualsAsHistogram`, `Plot_ResidualsVsTime`, `Plot_ResidualsVsObserved`, and `Plot_QQ`.
- Set `IndividualId` to `'*'`. In this way, all individuals are plotted in one panel. This is only available for goodness of fit plots, not for the time profile plots.

Do not forget to edit the columns `PlotCaptionAddon`, `TimeRange_firstApplication`, and `TimeRange_lastApplication`.

```{r plotconfig-2, echo=FALSE, eval = TRUE}
knitr::kable(dtPlots[5:11]  |>  dplyr::select(!any_of(c("TimeRange_h0_6", "TimeRange_h6_24"))))
```

Execute the `runPlot` call again to generate the plots. We take the newly generated plot names. The variable `plotList` contains all generated plots.

```{r plot2, eval = TRUE}
plotList <- runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  plotNames = c("selected_individuals-1", "selected_individuals-2", "all_individuals", "all_individuals-gof"),
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults
  )
)
```

Below the logarithmic time-profile for all individuals is displayed. See the Word document for the other plots.

```{r plot2figure, eval=TRUE,fig.height = 4.5, results='asis', fig.alt="Logarithmic time-profile plot for all individuals."}
print(plotList[["all_individuals-TP-log-total"]])
print(attr(plotList[["all_individuals-TP-log-total"]], "caption"))
```

## 3) Adding Random Populations and Use of Aggregated Data

In this section, we plot aggregated observed data against aggregated simulated random populations.

### Step 1: Data Aggregation

Since the observed data contains only individual-level records, we first compute group-level aggregates. Update your data import script as follows. The call to `updateDataGroupId()` is required to register the new aggregated groups in the `Reports.xlsx` configuration so that the plot function can find them.

To learn more about the aggregation options, see `?aggregateObservedDataGroups`.

```{r aggregateData, eval=FALSE}
dataObserved <- readObservedDataByDictionary(projectConfiguration = projectConfiguration)

# add aggregated  groups of data
dataObserved <- rbind(dataObserved,
  aggregateObservedDataGroups(
    dataObserved = dataObserved,
    groups = c("1234_iv", "1234_po")
  ),
  fill = TRUE
)
# For the manually added groups, the configuration sheet in Reports.xlsx has to be updated
updateDataGroupId(projectConfiguration, dataObserved)
```

### Step 2: Edit Population Configuration

Open the sheet `Demographics` in the `Populations.xlsx`.

In the example below, the population 'adults' is defined. As ontogeny, "CYP3A4" and "UGT1A4" are added.


```{r randompoulation-config, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$populationsFile,
  sheetName = "Demographics",
  skipDescriptionRow = FALSE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp[1])
```

### Step 3: Export Population

Go to the Export populations section of the workflow and uncomment lines to export random populations:

```{r randomPopulation, eval = FALSE}
exportRandomPopulations(
  projectConfiguration = projectConfiguration,
  populationNames = NULL,
  overwrite = FALSE
)
```

This will create a population based on the configuration table from the previous section. It is saved as a CSV file in `Models/Populations`.

### Step 4: Set Up and Run Scenario

Add two more scenarios to the `Scenarios` sheet:


```{r scenario-randompoulation, echo=FALSE, eval = TRUE}
knitr::kable(dtScenarios)
```

Since the simulation results from earlier sections are already cached on disk, set `scenarioNames = NULL` to run all scenarios including the new random-population ones. Previously cached results will be loaded from disk rather than re-simulated.

```{r simulate-random-poulations, eval=FALSE}
scenarioList <-
  createScenariosWrapped(
    projectConfiguration = projectConfiguration,
    scenarioNames = NULL
  )

scenarioResults <- runAndSaveScenarios(
  projectConfiguration = projectConfiguration,
  scenarioList = scenarioList,
  simulationRunOptions = ospsuite::SimulationRunOptions$new(
    numberOfCores = NULL,
    checkForNegativeValues = NULL,
    showProgress = TRUE
  )
)
```

### Step 5: Configure and Generate Plots

We will add the new scenarios to the plot configuration table using the helper function `addDefaultConfigForTimeProfilePlots`.

Then we open the configuration table and adjust the plot configurations.

In this example, we want both scenarios in one plot:

- Update the `PlotName` for both newly added scenarios to 'random_population'.
- As `DataGroupIds`, we can now use the groups added in Step 1: Data Aggregation.

Do not forget to edit the columns `TimeRange_firstApplication`, `TimeRange_lastApplication`, and add `LongName` entries in the sheet `Scenarios`.

```{r plotConfig-3, echo=FALSE, eval = TRUE}
knitr::kable(dtPlots[12:14]  |>  dplyr::select(!any_of(c("TimeRange_h0_6", "TimeRange_h6_24"))))
```

Execute the `runPlot` call again to generate the plots. 

```{r plot3, eval = TRUE}
plotList <- runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  plotNames = c("random_population"),
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults
  )
)
```

Below the logarithmic time-profile is displayed. See the Word document for the other plots.

```{r plot3figure, eval=TRUE,fig.height = 4, results='asis', fig.alt="Logarithmic time-profile plot for random populations."}
print(plotList[["random_population-TP-log-total"]])
print(attr(plotList[["random_population-TP-log-total"]], "caption"))
```

## 4) Adding Reference Scenarios

In this section, we overlay two scenarios on a single plot: `adults_po` as the primary scenario and `adults_iv` as the reference. The reference scenario is displayed with a distinct colour so the two curves can be compared directly.

### Configure and Generate Plots

- Copy the `adults_po` plot row and adjust `PlotName`.
- Set `ReferenceScenario` to `adults_iv`.
- Add legend labels for the two scenarios in the `ColorLegend` column.
- Use the `ScenarioName` column for a shared display name for both scenarios (this overrides the default `LongName` from the `Scenarios` sheet).

Do not forget to edit the columns `DataGroupIds`, `PlotCaptionAddon`, `TimeRange_firstApplication`, and `TimeRange_lastApplication`.

```{r plotConfig-4, echo=FALSE, eval = TRUE}
knitr::kable(dtPlots[15:16])
```

To adjust the reference colour, pass `referenceScaleVector` using the labels from the `colorLegend` column.

```{r plot4, eval = TRUE}
plotList <- runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  plotNames = c("reference"),
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults,
    referenceScaleVector = list("1mg iv simulation" = "grey")
  )
)
```

Below the logarithmic time-profile is displayed. See the Word document for the other plot.

```{r plot4figure, eval = TRUE, results='asis', fig.alt="Logarithmic time-profile plot comparing reference populations."}
print(plotList[["reference-TP-log-total"]])
print(attr(plotList[["reference-TP-log-total"]], "caption"))
```

## 5) Adding Plots with Different Time Ranges

Visualizing pharmacokinetic data across various time ranges can reveal different aspects of drug behavior. A typical example would be to show the first and last application in a multi-dosing simulation. As we have only a single dose in our example, we show the principle by splitting the time range into two slots: 0 - 6 hours and 6 to 24 hours.

### Step 1: Configure Time Ranges

In your plot configuration Excel file, in the sheet `TimeRange`, specify different time ranges for which you'd like to generate plots. This might involve adding new rows or adjusting existing settings in the `Reports.xlsx` file.

The `TimeRange` has the columns:

- `Tag`: identifier corresponds to the columns in the plot configuration sheet beginning with `TimeRange_`.
- `CaptionText`: Text added to the plot caption.
- `TimeLabel`: Label of the x-axis.
- `TimeShift`: An offset used to shift the time axis. If you want to display time after the last dose, your timeshift would be the time of the last dosing. (This is set automatically for the `firstApplication` and `lastApplication` tags if `TimeShift` is empty).

In the example below, a new row for the range between 0 and 6 hours and a row for the range between 6 and 24 hours have been added.


```{r config-TimeRange, echo=FALSE, eval = TRUE}
tmp <- xlsxReadData(
  wb = projectConfiguration$addOns$reportsFile,
  sheetName = "TimeRange",
  skipDescriptionRow = TRUE,
  emptyAsNA = FALSE,
  convertHeaders = FALSE
)

knitr::kable(tmp)
```

### Step 2: Configure and Generate Plots

Open the plot configuration table and adjust the plot configurations.

- Copy the plot line for the `one_individual` plot name to the end of the table and adjust `PlotName` for this line.
- Add new columns `TimeRange_h0_6` and `TimeRange_h6_24` for the newly created time ranges.
- Set the `TimeRange_h0_6` column to `c(0, 6)` and the `TimeRange_h6_24` column to `c(6, 24)`.
- Set the `TimeRange_total` column to `''`.

This configuration will produce one plot for each of the newly created time ranges.

We add another example where we display the time ranges in one plot:

- Copy the previous plot line and adjust the `PlotName`.
- Set the column `SplitPlotsPerTimeRange` to `FALSE` (0).
- Set the column `nFacetColumns` to 2. This way, in the final plot, we will have two columns for time ranges and three rows for outputs.
- Set the column `FacetScale` to `free_x`. All plots will have the same y-axis but different x-axes.


```{r plotconfig-5, echo=FALSE, eval = TRUE}
knitr::kable(dtPlots[17:19])
```

Execute the `runPlot` call again to generate the plots. The variable `plotList` contains all generated plots. In the example below, the input argument `xScaleArgs = list(limits = c(NA, NA))` is added. This argument is passed to the function `ospsuite.plots::plotTimeprofile` and overwrites the default time-profile x-limits, which is `c(0, NA)`.

```{r plot5, eval = TRUE}
plotList <- runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  plotNames = c("timeranges", "timeranges-split"),
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults,
    xScaleArgs = list(limits = c(NA, NA))
  )
)
```

Below the logarithmic time-profile for the plot that combines the time profiles is displayed. See the Word document for the other plots.

```{r plot5figure, eval = TRUE,fig.height=7, results='asis', fig.alt="Logarithmic time-profile plot showing combined time ranges."}
print(plotList[["timeranges-split-TP-log-allTimeRanges"]])
print(attr(plotList[["timeranges-split-TP-log-allTimeRanges"]], "caption"))
```

## 6) Report Creation

In this section, all plots from the previous sections are exported to disk and assembled into a single Word document. The workflow has two steps:

1. Call `runPlot()` without a `plotNames` argument so that every configured plot is generated and saved as an image and an `.Rmd` fragment.
2. Call `mergeRmds()` to concatenate the individual `.Rmd` fragments into one document, then `renderWord()` to produce the final `.docx`.

```{r plot6}
runPlot(
  nameOfplotFunction = "plotTimeProfiles",
  projectConfiguration = projectConfiguration,
  configTableSheet = "TimeProfiles",
  theme = runPlotTheme,
  inputs = list(
    dataObserved = dataObserved,
    scenarioResults = scenarioResults,
    referenceScaleVector = list("1mg iv simulation" = "grey"),
    xScaleArgs = list(limits = c(NA, NA))
  )
)
```

The first step is to run all plots without a `plotNames` filter so every configured plot is generated and saved. The second step merges the individual `.Rmd` fragments into a single document and renders it to Word.

```{r merge-Rmds}
mergeRmds(
  projectConfiguration = projectConfiguration,
  newName = "TutorialPlots",
  title = "Plots generated by tutorial",
  sourceRmds = c("TimeProfiles")
)

renderWord(fileName = file.path(projectConfiguration$outputFolder, "TutorialPlots.Rmd"))
```

## Conclusion

This tutorial covered the core time profile plotting workflow:

- Simulating a single individual and generating basic time profile and goodness of fit plots
- Building virtual twin and random populations and plotting aggregated results
- Comparing conditions using a reference scenario overlay
- Splitting plots across custom time-range windows
- Assembling all plots into a Word report

For next steps, see the other vignettes in this package:

- **[Data Import by Dictionary](DataImport_by_Dictionary.html)** — detailed data import and dictionary configuration
- **[Population](Population.html)** — population configuration and export options
- **[Workflow](Workflow.html)** — the full workflow script structure and options
