---
title: "Workflow Setup and Execution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  error = FALSE,
  message = FALSE,
  eval = FALSE,
  comment = "#>"
)
```

```{r setup,echo=FALSE, eval = TRUE}
library(ospsuite.reportingframework)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Typical Package Workflow

This vignette explains how to set up and run a reporting workflow with `{ospsuite.reportingframework}`.

### Accessing the Workflow Template

To access the default workflow template provided by the package, you can use the following function:

```{r, echo=TRUE, eval=FALSE}
openWorkflowTemplate()
```

This function opens the workflow template as a new document in your RStudio environment, allowing you to customize it for your specific needs. For detailed information on what each function in the workflow does, please refer to the help functionality in R.

#### Access via RStudio Addins

You can also access the workflow template through an RStudio Addin. To use the Addins, simply navigate to the Addins menu in RStudio and select "Open Template Workflow." This will execute the `openWorkflowTemplate()` function and open the template in a new document.

#### Customizing the Workflow Template

By default, the workflow template is loaded from the package's internal templates. If you wish to customize this template, follow these steps:

1. Create your own workflow template script and save it with the name `template_workflow.R`.
2. Place this file in your designated template directory.
3. Set the path to your custom template by using the following option:

```{r, echo=TRUE, eval=FALSE}
options(OSPSuite.RF.PathForWorkflowTemplate = "myTemplateDirectory")
```

After setting this option, calling `openWorkflowTemplate()` or using the Addins will open your customized template instead of the default one. You can save this option in your ".Rprofile" file (`file.path(Sys.getenv("HOME"), ".Rprofile"`)) to set it automatically.

## Project Initialization

The first step in the workflow is to initialize the project. This includes:

- Loading necessary libraries.
- Setting the workflow options.
- Setting up the project directory and configuration files using the project structure of `{esqlabsR}`.
- Creating or updating `.gitignore` files in configuration folders so `*.xlsx` configuration files are ignored by Git.
- Initializing the log file.

Set workflow behavior with `setWorkflowOptions(isValidRun)`. This controls whether exploratory outputs are allowed (`isValidRun = FALSE`) or a QC-passed run is enforced (`isValidRun = TRUE`). If `isValidRun` is omitted, `getQCpassedEnvironmentVariable()` reads `QCpassed` from the environment.

```{r, echo=TRUE, eval=FALSE}
# Exploratory run: enable watermarks and allow helper functions
setWorkflowOptions(isValidRun = FALSE)

# QC-passed run: disable watermarks and enforce strict execution
setWorkflowOptions(isValidRun = TRUE)
```

Create and load the project configuration from the project workbook path:

```{r, echo=TRUE, eval=FALSE}
projectConfiguration <- createProjectConfiguration(
  path = file.path(configurationDirectory, "ProjectConfiguration.xlsx"),
  ignoreVersionCheck = FALSE
)
```

`ProjectConfiguration.xlsx` stores the package versions used to create the project: `esqlabsRVersion` (managed by `{esqlabsR}`) and `ospsuiteReportingFrameworkVersion` (managed by this package).
When you load a configuration, both versions are checked against installed packages; mismatches prompt for update in interactive sessions and stop execution in non-interactive runs (for example CI) until resolved. New projects created with `initProject()` receive the current framework version automatically, and `projectConfiguration$save()` updates stored version values.

### Configuration in Git: Use JSON Snapshots

Configuration workbooks (`*.xlsx`) are generated working files and are ignored by Git in configuration folders. The versioned source of truth is the JSON snapshot created from the configuration.

Save current configuration to JSON:

```{r, echo=TRUE, eval=FALSE}
projectConfiguration <- createProjectConfiguration(
  path = file.path(configurationDirectory, "ProjectConfiguration.xlsx")
)

snapshotProjectConfigurationRF(
  projectConfiguration,
  outputDir = configurationDirectory
)
```

This writes `ProjectConfiguration.json` in `configurationDirectory`. Commit this JSON file to Git.

Restore configuration from JSON (for example on another machine or in CI):

```{r, echo=TRUE, eval=FALSE}
projectConfiguration <- restoreProjectConfigurationRF(
  snapshotFilePath = file.path(configurationDirectory, "ProjectConfiguration.json"),
  outputDir = configurationDirectory
)
```

## Data Import

After initialization, the first step should always be data import. The input format for data is CSV files, which can be imported using a dictionary. The imported data types include individual and aggregated time profile data as well as individual and aggregated PK parameter data.

For more information, see **[Data Import by Dictionary](DataImport_by_Dictionary.html)**.

## Setting Up and Running Simulations Including PK Parameter Calculation

Simulations are set up and run using the `{esqlabsR}` package. For more information and detailed documentation, users can visit the [esqlabsR Documentation](https://esqlabs.github.io/esqlabsR/).

The `{ospsuite.reportingframework}` package provides additional functionality to:

- Export random and virtual twin populations (see the **[Population](Population.html)**).
- Calculate PK Parameters (see the **[PK Parameters](PK_Parameter.html)**).

## Plot and Report Generation

The core functionality of this package lies in plot and report creation. It includes various plotting functions that generate visualizations based on the analyses performed. 

Each plot function is designed to create an R Markdown (`.Rmd`) file, which contains the code and output for the respective plot. This allows for easy documentation and sharing of results.

The package provides a collection of plot functions and interfaces to adjust plots and utilize customized plot functions. 

For more information, see the **[Plot and Report Generation](Plot_and_Report_Generation.html)**.
